Natural Stories LM analysis

### 2020-04-30 Analysis/nat_stories_lms.Rmd
cleaned up with only current analysis

Surprisals are measured in bits (TODO: double check with Jenn about these being in bits, model specs/cites!)

@@

The model formula we use is rt ~ surp + length * freq + past_surp + past_length * past_freq, so we're allowing frequency and length to interact and including the same terms for the current and past word.

@@

## Linear models

These use centered predictors.

Using pre-error data only.

```{r lm-2}

lm_ngram_model_1 <- lm(rt ~ ngram_center + freq_center * length_center+ past_c_ngram + past_c_freq*past_c_length, data=labelled_pre_error)

lm_txl_model_1 <- lm(rt ~ txl_center + freq_center * length_center+ past_c_txl + past_c_freq*past_c_length, data=labelled_pre_error)

lm_grnn_model_1 <- lm(rt ~ grnn_center + freq_center * length_center+ past_c_grnn + past_c_freq*past_c_length, data=labelled_pre_error)

```{r lm_summary-2}

summary(lm_ngram_model_1)

summary(lm_txl_model_1)

summary(lm_grnn_model_1)

```
Including post-error data.

```{r lm-3}

lm_ngram_model_2 <- lm(rt ~ ngram_center + freq_center * length_center+ past_c_ngram + past_c_freq*past_c_length, data=labelled_post_error)

lm_txl_model_2 <- lm(rt ~ txl_center + freq_center * length_center+ past_c_txl + past_c_freq*past_c_length, data=labelled_post_error)

lm_grnn_model_2 <- lm(rt ~ grnn_center + freq_center * length_center+ past_c_grnn + past_c_freq*past_c_length, data=labelled_post_error)

```{r lm_summary-3}

summary(lm_ngram_model_2)

summary(lm_txl_model_2)

summary(lm_grnn_model_2)

@@

We include a by-subject effect for everything. Technically speaking, we do have repeated measures at the word level, but there are 10K or so words and at most a handful of copies of each word, so that's likely to get us not very much at the cost of making the model rather intractable (and so I'm not going to deal with it, at least not right now).

@@

 brm_ngram_model <- brm(rt ~ ngram_center + freq_center * length_center +
                         past_c_ngram + past_c_freq*past_c_length +
                         (ngram_center + freq_center * length_center +
                            past_c_ngram + past_c_freq*past_c_length|subject), data=labelled_pre_error,
                       file="brm_ngram_pre_center", prior=priors, control=list(adapt_delta=.95))

brm_grnn_model <- brm(rt ~ grnn_center + freq_center * length_center +
                         past_c_grnn + past_c_freq*past_c_length +
                         (grnn_center + freq_center * length_center +
                            past_c_grnn + past_c_freq*past_c_length|subject), data=labelled_pre_error,
                       file="brm_grnn_pre_center", prior=priors, control=list(adapt_delta=.95))

brm_txl_model <- brm(rt ~ txl_center + freq_center * length_center +
                         past_c_txl + past_c_freq*past_c_length +
                         (txl_center + freq_center * length_center +
                            past_c_txl + past_c_freq*past_c_length|subject), data=labelled_pre_error,
                       file="brm_txl_pre_center", prior=priors, control=list(adapt_delta=.95))

@@

  stats
}
```
```{r}
summary(brm_txl_model)
summary(brm_grnn_model)
summary(brm_ngram_model)
```

```{r}
a <- show_summary(brm_txl_model) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
b <- show_summary(brm_grnn_model) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
c <- show_summary(brm_ngram_model) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))

@@

@@

>>>>

Surprisals are measured in bits.

@@

The model formula we use is rt ~ surp* length +  freq * length + past_surp * past_length + past_freq * past_length.

@@

```
```{r}

lab_summary <- select_labs %>% summarize_at(c("length","freq","txl_surp","grnn_surp","ngram_surp"), list(mean=mean, stdev=sd, min=min, max=max)) %>% 
  pivot_longer(everything()) %>% arrange(name)

ggplot(select_labs, aes(length))+geom_bar()

@@

We include a by-subject effect for everything, and a by_word random intercept (full mixed effects).

@@

### On pre-error data only 
```{r}
brm_ngram_interact <- brm(rt ~ ngram_center* length_center + freq_center * length_center + 
                         past_c_ngram*past_c_length + past_c_freq*past_c_length + 
                         (ngram_center* length_center + freq_center * length_center + 
                            past_c_ngram *past_c_length+ past_c_freq*past_c_length|subject)+
                         (1|Word_ID), data=labelled_pre_error, 
                       file="brm_ngram_interact", prior=priors, control=list(adapt_delta=.95))

brm_grnn_interact <- brm(rt ~ grnn_center * length_center+ freq_center * length_center + 
                         past_c_grnn *past_c_length+ past_c_freq*past_c_length + 
                         (grnn_center* length_center + freq_center * length_center + 
                            past_c_grnn*past_c_length + past_c_freq*past_c_length|subject)+
                        (1|Word_ID), data=labelled_pre_error, 
                       file="brm_grnn_interact", prior=priors, control=list(adapt_delta=.95))

brm_txl_interact <- brm(rt ~ txl_center* length_center + freq_center * length_center + 
                         past_c_txl*past_c_length + past_c_freq*past_c_length + 
                         (txl_center* length_center + freq_center * length_center + 
                            past_c_txl *past_c_length+ past_c_freq*past_c_length|subject)+
                       (1|Word_ID), data=labelled_pre_error, 
                       file="brm_txl_interact", prior=priors, control=list(adapt_delta=.95))

@@

  stats
}
a <- show_summary(brm_txl_interact) %>% mutate(model="TXL") %>% mutate(Term=str_replace(Term,"txl","surp"))
b <- show_summary(brm_grnn_interact) %>% mutate(model="GRNN") %>% mutate(Term=str_replace(Term,"grnn","surp"))
c <- show_summary(brm_ngram_interact) %>% mutate(model="5-gram") %>% mutate(Term=str_replace(Term,"ngram","surp"))

@@

### On post-error as well

```{r}
brm_ngram_interact_post <- brm(rt ~ ngram_center* length_center + freq_center * length_center + 
                         past_c_ngram*past_c_length + past_c_freq*past_c_length + 
                         (ngram_center* length_center + freq_center * length_center + 
                            past_c_ngram *past_c_length+ past_c_freq*past_c_length|subject)+
                         (1|Word_ID), data=labelled_post_error, 
                       file="brm_ngram_interact_post", prior=priors, control=list(adapt_delta=.95))

brm_grnn_interact_post <- brm(rt ~ grnn_center * length_center+ freq_center * length_center + 
                         past_c_grnn *past_c_length+ past_c_freq*past_c_length + 
                         (grnn_center* length_center + freq_center * length_center + 
                            past_c_grnn*past_c_length + past_c_freq*past_c_length|subject)+
                        (1|Word_ID), data=labelled_post_error, 
                       file="brm_grnn_interact_post", prior=priors, control=list(adapt_delta=.95))

brm_txl_interact_post <- brm(rt ~ txl_center* length_center + freq_center * length_center + 
                         past_c_txl*past_c_length + past_c_freq*past_c_length + 
                         (txl_center* length_center + freq_center * length_center + 
                            past_c_txl *past_c_length+ past_c_freq*past_c_length|subject)+
                       (1|Word_ID), data=labelled_post_error, 
                       file="brm_txl_interact_post", prior=priors, control=list(adapt_delta=.95))

```
```{r}
show_summary_2 <- function(model){
  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()

  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
    group_by(.variable) %>% 
    summarize(pct_above_0=mean(above_0)) %>% 
    mutate(`P-value equivalent` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
    left_join(intervals, by=".variable") %>% 
    mutate(lower=round(.lower, digits=2),
           upper=round(.upper, digits=2),
           `Credible Interval`=str_c("[",lower,", ", upper,"]"),
           Term=str_sub(.variable, 3, -1),
           Estimate=round(.value, digits=2)) %>% 
    select(Term, Estimate, `Credible Interval`, `P-value equivalent`)

  stats
}
```

### Summaries

```{r}
summary(brm_ngram_interact_post)
summary(brm_ngram_interact)

summary(brm_grnn_interact_post)
summary(brm_grnn_interact)

summary(brm_txl_interact_post)
summary(brm_txl_interact)
```

@@

# TODOs for future

In terms of surprisal analyses, the possible next steps are doing GAM analysis more robustly (with mixed effects, presenting results from the models) and doing DTSR.

We could also try looking at the specific syntactic features that are in the dataset (the point of natural stories was to have a bunch of idioms and weird syntax); we could maybe analyse those?
